---
title: "Higher Dimensional Continuous Data"
author: "Zihan Guo"
date: "November 2, 2016"
output: 
  html_document:
    toc:  true
    toc_float:  true
    code_folding:  hide
---

#  Higher Dimensional Continuous Data

***
***

#  Task 0

**Organization, Themes, and HTML Output**

Wickham R Style

```{r, echo=TRUE, message=FALSE, warning=FALSE}

# prepare data, library and theme. 
library(ggplot2)
library(dplyr)
library(data.table)
library(devtools)
library(MASS)

zihang_315_theme <- theme(
    axis.text = element_text(size = 12, family = 'mono', colour = 'black'),
    axis.title = element_text(size = 12, family = 'mono', colour = 'black'),
    legend.title = element_text(size = 10, family = 'mono', colour = 'black'),
    plot.title = element_text(size = 14,  family = 'mono', colour = 'black'),
    legend.text = element_text(size = 8, family = 'mono', colour = 'black'),
    legend.key = element_rect(fill = "white"),
    legend.background = element_rect(fill = "white"),
    panel.grid.major = element_line(colour = "grey"),
    panel.grid.minor = element_blank(),
    panel.background = element_rect(fill = "white")
  )

data(Cars93)
olive <- fread("https://raw.githubusercontent.com/sventura/315-code-and-datasets/master/data/olive_oil.csv")

```


***
***


#  Task 1 {.tabset}

##  Part (a)

```{r, echo=TRUE, message=FALSE, warning=FALSE}

# Problem 1 a : Using textual annotations 
ggplot(data = Cars93, aes(x = Fuel.tank.capacity, y = MPG.city)) + 
  geom_text(aes(label = Cars93$Model)) + 
  zihang_315_theme + 
  ggtitle("Scatterplor of City MPG and Fuel Tank Capacity") + 
  labs(x = "Fuel tank capacity (US gallons)", y = "City MPG (miles per US gallon)")

```


##  Part (b)

```{r, echo=TRUE, message=FALSE, warning=FALSE}

cbbPalette <- c("#000000", "#E69F00", "#56B4E9", "#009E73", "#F0E442", "#0072B2", "#D55E00", "#CC79A7")

# Problem 1 b : Changing Angle of Text
# Q: how to change legand title? and to non-defualt color scheme? 
ggplot(data = Cars93, aes(x = Fuel.tank.capacity, y = MPG.city)) + 
  geom_text(aes(color = factor(Cars93$Type), label = Cars93$Model, angle = 30)) +
  labs(colour="Type of Car") + 
  zihang_315_theme + 
  ggtitle("Scatterplor of City MPG and Fuel Tank Capacity") + 
  labs(x = "Fuel tank capacity (US gallons)", y = "City MPG (miles per US gallon)") + 
  scale_colour_manual(values=cbbPalette)
```

##  Part (c)

```{r, echo=TRUE, message=FALSE, warning=FALSE}

# Problem 1 c : Size and Font Family

ggplot(Cars93, aes(x = Fuel.tank.capacity, y = MPG.city), label = Model) + 
  geom_text(aes(color = factor(Type), 
                label = Model, angle = 30,  
                size = RPM, 
                fontface = 2)) + 
  scale_size("RPM (revs per minute \nat maximum horsepower)") + 
  labs(colour="Type of Car") + 
  # scale_color_brewer(palette="Dark2") + 
  zihang_315_theme + 
  ggtitle("Scatterplor of City MPG and \nFuel Tank Capacity") + 
  labs(x = "Fuel tank capacity (US gallons)", y = "City MPG (miles per US gallon)") +
  scale_colour_manual(values=cbbPalette)

```


***
***


#  Task 2 {.tabset}


**More Text Annotations on Graphs**

##  Part (a)

```{r, echo=TRUE, message=FALSE, warning=FALSE}

number_ticks <- function(n) {function(limits) pretty(limits, n)}
# Citation: 
# http://stackoverflow.com/questions/11335836/increase-number-of-axis-ticks-in-ggplot2 

# Problem 2 a. 
ggplot(data = olive, aes(region)) + 
      geom_bar() + 
      labs(y = "Count", x = "Region") + 
      geom_text(stat = 'count', aes(label = ..count..), vjust = -0.3) + 
      ggtitle("Distribution of Olive by Region") + 
      zihang_315_theme + 
      scale_y_continuous(breaks=number_ticks(10)) + 
      scale_x_continuous(breaks=number_ticks(10))

```

##  Part (b)

```{r, warning = F, message = F}

#  Problem 2 b: 
height <- table(olive$area)

ggplot(data = olive, aes(area)) + 
      geom_bar(fill = "light grey") + 
      labs(y = "Count", x = "Area") + 
      geom_text(stat = 'count', aes(label = ..count..), y = height/2) + 
      ggtitle("Distribution of Olive by Area") + 
      zihang_315_theme + 
      scale_y_continuous(breaks=number_ticks(3)) + 
      scale_x_continuous(breaks=number_ticks(3))

```

##  Part (c)

```{r, warning = F, message = F}

#  Your code for part (c) here
ggplot(data = olive, aes(area)) + 
      geom_bar(fill = "light grey") + 
      labs(y = "Count", x = "Area") + 
      geom_text(stat = 'count', 
                aes(label = scales::percent((..count..)/sum(..count..))), 
                y = height/2) + 
      ggtitle("Distribution of Olive by Area") + 
      zihang_315_theme + 
      scale_y_continuous(breaks=number_ticks(3)) + 
      scale_x_continuous(breaks=number_ticks(3))

```

***
***



#  Task 3  {.tabset}

**2D-KDEs with Contour Plots and Adjusted Bandwidths**

##  Part A

```{r, message=FALSE, warning=FALSE}


olive$area <- factor(olive$area)
olive$region <- factor(olive$region)

olive_cont <- subset(olive, 
                     select = c(palmitic, palmitoleic, stearic, oleic, linoleic,
                                linolenic, arachidic, eicosenoic))
olive_cont_scale <- scale(x = olive_cont)
dist_olive <- dist(x = olive_cont_scale)
olive_mds <- cmdscale(dist_olive, k = 2)
olive_mds <- as.data.frame(olive_mds)
colnames(olive_mds) <- c("mds_coordinate_1", "mds_coordinate_2")
olive_mds <- mutate(olive_mds, 
                    area = as.factor(olive$area), 
                    region = as.factor(olive$region))


```

##  Part B

```{r, message=FALSE, warning=FALSE}

ggplot(data = olive_mds) + 
  geom_density2d(aes(x = mds_coordinate_2, y = mds_coordinate_1)) + 
  geom_point(aes(y = mds_coordinate_2, x = mds_coordinate_1)) + zihang_315_theme + 
  labs(x = "Coordinate 1", y = "Coordinate 2") +
  ggtitle("Contour Plot of Olive MDS")

```


##  Part C

```{r, message=FALSE, warning=FALSE}

ggplot(data = olive_mds) + 
  geom_density2d(aes(x = mds_coordinate_2, y = mds_coordinate_1)) + 
  geom_point(aes(x = mds_coordinate_2, y = mds_coordinate_1, 
                 color = area, shape = region), size = 3) + 
  ggtitle("Countour Map of MDS Coordinates") + 
  scale_shape_manual(values = as.character(sort(unique(olive_mds$region)))) + 
  labs(x = "Coordinate 1", y = "Coordinate 2", color = "Area(1, 2, 3)") + 
  zihang_315_theme

```


##  Part D

```{r, message=FALSE, warning=FALSE}

ggplot(data = olive_mds) + 
  geom_density2d(h = c(1,1), aes(x = mds_coordinate_2, y = mds_coordinate_1)) + 
  geom_point(aes(x = mds_coordinate_2, y = mds_coordinate_1, 
                 color = area, shape = region), size = 1) + 
  ggtitle("Countour Map of MDS Coordinates")  + zihang_315_theme + 
  labs(x = "Coordinate 1", y = "Coordinate 2", color = "Area(1, 2, 3)")

```

##  Part E

```{r, message=FALSE, warning=FALSE}

ggplot(data = olive_mds) + 
  geom_density2d(h = c(5, 5), aes(x = mds_coordinate_2, y = mds_coordinate_1)) + 
  geom_point(aes(x = mds_coordinate_2, y = mds_coordinate_1, 
                 color = area, shape = region), size = 1) + 
  ggtitle("Countour Map of MDS Coordinates") + 
  labs(x = "Coordinate 1", y = "Coordinate 2", color = "Area(1, 2, 3)") + zihang_315_theme

```

##  Part F

##  Part G

```{r, message=FALSE, warning=FALSE}

ggplot(data = olive_mds) + 
  geom_density2d(h = c(2, 1), aes(x = mds_coordinate_1, y = mds_coordinate_2)) + 
  geom_point(aes(x = mds_coordinate_1, y = mds_coordinate_2, 
                 color = area, shape = region), size = 3) + 
  ggtitle("Countour Map of MDS Coordinates") + 
  scale_shape_manual(values = as.character(sort(unique(olive_mds$region)))) + 
  scale_colour_manual(values = c("blue", "light blue", "yellow"))

```


***
***


#  Task 4 {.tabset}

**2D-KDEs with Heat Maps and Three-Color Gradients**

## Part A

```{r, message=FALSE, warning=FALSE}

library(scales)
ggplot(data = olive_mds) + 
  stat_density2d(h = c(2, 1), aes(fill = ..density.., 
                     x = mds_coordinate_1, y = mds_coordinate_2), 
                 geom = "tile", contour = F) + 
  scale_fill_gradient2(mid= "orange", high ="red", low="white", midpoint = 0.05) + 
  theme_classic() + 
  zihang_315_theme + 
  geom_point(aes(x = mds_coordinate_1, y = mds_coordinate_2, 
                 color = area, shape = region), size = 3) +
  scale_shape_manual(values = as.character(sort(unique(olive_mds$region)))) + 
  scale_colour_manual(values = c("blue", "light blue", "yellow")) + 
  labs(color = "Area (1, 2, 3)", x = "Coordinate 1", y = "Coordinate 2") + 
  ggtitle("Heat Map of Two Dimensional \n DKEs")

```


***
***



#  Task 5 {.tabset}

**Hierarchical Clustering and Dendrograms**

## Part A

```{r, echo=TRUE, message=FALSE, warning=FALSE}

olive_cont <- subset(olive, 
                     select = c(palmitic, palmitoleic, stearic, oleic, linoleic,
                                linolenic, arachidic, eicosenoic))
olive_cont_scale <- scale(x = olive_cont)
dist_olive <- dist(x = olive_cont_scale)

```

## Part B

```{r, echo=TRUE, message=FALSE, warning=FALSE}

hc_olive_complete <- hclust(dist_olive, method = "complete")
names(hc_olive_complete)
hc_olive_complete$method

```


## Part C


```{r, echo=TRUE, message=FALSE, warning=FALSE}

plot(hc_olive_complete)

```

## Part D

## Part E


## Part F

```{r, echo=TRUE, message=FALSE, warning=FALSE}

labels_complete_2 <- cutree(hc_olive_complete, k = 2)
typeof(labels_complete_2)
length(labels_complete_2)

```

## Part G

```{r, echo=TRUE, message=FALSE, warning=FALSE}

table(labels_complete_2) / nrow(olive)

```


## Part H 

```{r, echo=TRUE, message=FALSE, warning=FALSE}

labels_complete_3 <- cutree(hc_olive_complete, k = 3)
olive_cont <- subset(olive, 
                     select = c(palmitic, palmitoleic, stearic, oleic, linoleic,
                                linolenic, arachidic, eicosenoic))
olive_cont_scale <- scale(x = olive_cont)
dist_olive <- dist(x = olive_cont_scale)
olive_mds <- cmdscale(dist_olive, k = 2)
olive_mds <- as.data.frame(olive_mds)
colnames(olive_mds) <- c("mds_coordinate_1", "mds_coordinate_2")
olive_mds <- mutate(olive_mds, 
                    area = as.factor(olive$area), 
                    region = as.factor(olive$region),
                    labels_complete_3 = as.factor(labels_complete_3))


```


```{r, echo=TRUE, message=FALSE, warning=FALSE}

ggplot(data = olive_mds) + 
  geom_point(aes(color = area, shape = labels_complete_3,
                 x = mds_coordinate_1, y = mds_coordinate_2), size = 5) + 
  scale_shape_manual(values = as.character(sort(unique(olive_mds$area)))) + 
  ggtitle("Scatterplot of Coordinate 1 and Coordinate 2") + zihang_315_theme + 
  xlab("Coordinate 1") + ylab("Coordinate 2") + 
  labs(colour="Area (1, 2, 3)") 

```


## Part I

```{r, echo=TRUE, message=FALSE, warning=FALSE}

library(dendextend)
library(dplyr)
library(MASS)
library(ggplot2)
data(Cars93)
colorblind_palette <- c("#999999", "#E69F00", "#56B4E9", "#009E73", "#F0E442", "#0072B2", "#D55E00", "#CC79A7")

olive_cont <- subset(olive, 
                     select = c(palmitic, palmitoleic, stearic, oleic, linoleic,
                                linolenic, arachidic, eicosenoic))

olive_mds$regions <- factor(olive_mds$region,
                            levels = c(1, 2, 3, 4, 5, 6, 7, 8, 9),
                            labels = c('northern Apulia', 'southern Apulia', 
                                       'Calabria', 'Sicily', 'inland Sardinia',
                                       'coast Sardinia', 'eastern',
                                       'western Liguria', 'Umbria'))

get_colors <- function(x, palette = colorblind_palette) palette[match(x, unique(x))]

dend <- olive_cont %>% scale %>% 
  dist %>% hclust %>% as.dendrogram %>% 
  set("labels", olive_mds$regions) %>% 
  set("labels_col", get_colors(olive_mds$labels_complete_3)) 

ggd1 <- as.ggdend(dend)
ggplot(ggd1, horiz = TRUE, theme = zihang_315_theme) +
  xlab("Dendrogram of Olive MDS Region Name") + 
  ylab("Pairwise Euclidean Distance")

```

## Part J 

```{r, echo=TRUE, message=FALSE, warning=FALSE}

dend <- olive_cont %>% scale %>% 
  dist %>% hclust %>% as.dendrogram %>% 
  set("branches_k_color", olive_mds$regions) %>% 
  set("labels_col", get_colors(olive_mds$labels_complete_3))

ggd1 <- as.ggdend(dend)
ggplot(ggd1, horiz = TRUE, theme = zihang_315_theme) +
  xlab("Dendrogram of Olive MDS Region Name") + 
  ylab("Pairwise Euclidean Distance")

```

***
***

#  Task 6

**Beyond Default Pairs Plots**

```{r, echo=TRUE, message=FALSE, warning=FALSE}

library(GGally)
olive_sub <- subset(olive, 
                    select = c(palmitic, stearic, oleic))

olive_sub <- mutate(olive_sub, 
                    area = as.factor(olive$area), 
                    region = as.factor(olive$region))
                    
ggpairs(data = olive_sub,
        upper = list(continuous = "density",
                     discrete = "ratio",
                     combo = "facetdensity"),
        lower = list(continuous = "cor",
                     discrete = "ratio",
                     combo = "box"),
        title = "Pairplot")

```



***
***